---
title: "plot CRISPR sgRNA"
author: "Karen Chu"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library

```{r lib}
library(openxlsx)
library(dplyr)
library(Biostrings)
library(stringr)
library(ggplot2)

library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(BSgenome.Hsapiens.UCSC.hg19)
```

Work dir

```{r work dir}
figures.dir <- "~/Desktop/Nynrin/figures/"
output.dir <- "~/Desktop/Nynrin/output/"
```

Import data

```{r data}
df.original <- read.csv("~/Desktop/Nynrin/data/NYNRIN_D3_vs_D17_All_Karen.csv")
df <- df.original %>% dplyr::select(c("logFC", "SEQ"))
rownames(df) <- df.original$NAME
```

Get genomic coords of gene of interest

```{r genomic coords}
txdb.genes = genes(TxDb.Hsapiens.UCSC.hg19.knownGene)

# Get genomic coord of gene
gene.coords <- subset(txdb.genes, gene_id==57523) 
```

Get sequence of gene of interest.

```{r seq}
genome <- BSgenome.Hsapiens.UCSC.hg19

# Obtain sequence
gene.seq <- Biostrings::getSeq(genome, gene.coords) # second argument needs grange object with coordinates'

# Obtain sequence as string
gene.seq.df <- as.data.frame(gene.seq) 
```

Get reverse complement of sequence

```{r reverse}
reverse.gene.seq <- reverseComplement(gene.seq)
reverse.gene.seq.df <- as.data.frame(reverse.gene.seq) 
```

Remove controls

```{r controls}
df <- subset(df, !(grepl("Safe|Control", rownames(df))) )
```

Obtain start and end position of sgRNA sequence in gene sequence on + strand

```{r index}
setwd(output.dir)

df.positive.strand <- df

df.positive.strand$start.pos <- NA
df.positive.strand$end.pos <- NA

for (i in 1:nrow(df)) {
  
  sgRNA.index <- gregexpr(df.positive.strand$SEQ[i], gene.seq.df) # Returns positions of every match in a string
  start.pos <- unlist(sgRNA.index)
  
  if ( start.pos == -1 ) {
    
    next
    
  } else { 
    
    end.pos <- start.pos + attr(sgRNA.index[[1]],'match.length') # Add start position with length of match
  
    # Add positions to dataframe
    df.positive.strand$start.pos[i] <- start(gene.coords) + start.pos - 1
    df.positive.strand$end.pos[i] <- start(gene.coords) + end.pos - 1
    
    }
  
}

write.csv(df.positive.strand, "Nynrin_sgRNA_matched_index_positive_strand.csv")
```


Obtain start and end position of sgRNA sequence in gene sequence on - strand


```{r index}
setwd(output.dir)

df.negative.strand <- df

df.negative.strand$start.pos <- NA
df.negative.strand$end.pos <- NA

for (i in 1:nrow(df)) {
  
  sgRNA.index <- gregexpr(df.negative.strand$SEQ[i], reverse.gene.seq.df) # Returns positions of every match in a string
  start.pos <- unlist(sgRNA.index)
  if ( start.pos == -1 ) {

    next
    
  } else { 
    
    # Add positions to dataframe
    # Add or subtract 1 because of R indexing
    df.negative.strand$start.pos[i] <- end(gene.coords) - start.pos - attr(sgRNA.index[[1]],'match.length') + 1
    df.negative.strand$end.pos[i] <- end(gene.coords) - start.pos + 1
    
    }
  
  
}

write.csv(df.negative.strand, "Nynrin_sgRNA_matched_index_negative_strand.csv")
```

Combine positions of sgRNA from + and - strand together.

```{r combine}
setwd(output.dir)

# Obtain index of sgRNA matching to opposite strand
missing.value.index.positive.strand <- which( is.na(df.positive.strand$start.pos))
missing.value.index.negative.strand <- which( is.na(df.negative.strand$start.pos))

# Create final dataframe with start and end positions from + and - strand
df.final <- df.positive.strand
df.final$start.pos[missing.value.index.positive.strand] <- df.negative.strand$start.pos[missing.value.index.positive.strand]
df.final$end.pos[missing.value.index.positive.strand] <- df.negative.strand$end.pos[missing.value.index.positive.strand]

write.csv(df.final, "Nynrin_sgRNA_matched_index_final_plot_input.csv")
```

Get exon coordinates of gene of interest.

```{r exon}
setwd(output.dir)

# Get genomic coordinates of transcripts
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene
transcripts <- transcriptsBy(txdb, by="gene")
gene.transcripts.index <- subjectHits( findOverlaps(gene.coords, transcripts) )
gene.transcripts.final <- transcripts[gene.transcripts.index]

gene.transcripts.final.df <- as.data.frame(gene.transcripts.final)
tx_names <- gene.transcripts.final.df$tx_name
exons <- exonsBy(txdb, "tx", use.names=TRUE)[tx_names]

write.csv(exons[[1]], "Nynrin_exon_coordinates.csv")
```

Calculate exon positions relative to start position

```{r position relative to start}
# Calculate positions relative to if start of gene is 0 coordinate

calculate.relative.position <- function(exons.df) {
  
  isoform<- as.data.frame(exons.df)
  isoform$start.relative <- isoform$start - start(gene.coords )
  isoform$end.relative <- isoform$end - start(gene.coords)
  
  return(isoform)
  
}

isoform.one <- calculate.relative.position(exons[[1]])
```

Plot gene with sgRNA logfc
Plot negative and positive sgRNA logfc separately for visual aesthetics.

```{r mid point}
setwd(figures.dir)

# Calculate mid-point between start and end pos
df.final$mid.point <- df.final$start.pos + ( (df.final$end.pos - df.final$start.pos) /2)

# Change negative logfc to positive
negative.logfc.only <- subset(df.final, logFC < 0)
negative.logfc.only$logFC <- negative.logfc.only$logFC*-1
positive.logfc.only <- subset(df.final, logFC >= 0)

plot.gene <- function(df.input, isoform.input, pos.or.neg.logfc) {
  
  png(paste0("Nynrin_sgRNA_", pos.or.neg.logfc, ".png"), 1500, 500)
  p<-ggplot(data=df.input, aes(x=mid.point, y=logFC)) +
    geom_bar(stat="identity", fill="darkred", width=50) +
    geom_segment(x=start(gene.coords), y= -0.4,
                 xend=end(gene.coords), 
                 yend= -0.4,
                 size=3, color="black") +
    theme_minimal() +
    #xlab("\nGene Coordinates") + 
    ggtitle(paste0("Nynrin: ", pos.or.neg.logfc, " sgRNAs\n")) +
    ylab("Log Fold Change\n") +
    xlim( start(gene.coords), end(gene.coords) ) +
    ylim(0,max(c(negative.logfc.only$logFC, positive.logfc.only$logFC))+0.5) + 
    theme(plot.title = element_text(size=22),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y   = element_text(size=20),
          #axis.text.x   = element_text(size=20),
          axis.title.y  = element_text(size=20),
          #axis.title.x  = element_text(size=20),
          panel.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.border = element_rect(colour = "black", fill=NA, size=2)) +
    coord_cartesian(clip="off") +
    theme(plot.margin=unit(c(2,2,2,2),"cm"))

  # Add geom segments with for loop
  for (i in 1:nrow(isoform.input)) {
     p <- p + geom_segment(x=isoform.input$start[i], y= -0.4, 
                           xend=isoform.input$end[i], yend= -0.4,
                           size=5,
                           color="gray")
  }
  
  print(p)
  
  dev.off()
  
}

plot.gene(negative.logfc.only, isoform.one, "Negative.Logfc")
plot.gene(positive.logfc.only, isoform.one, "Positive.Logfc")
```

Positive and negative logfc on same plot

```{r combined plot}
setwd(figures.dir)

# Calculate mid-point between start and end pos
df.final$mid.point <- df.final$start.pos + ( (df.final$end.pos - df.final$start.pos) /2)

plot.gene <- function(df.input, isoform.input, isoform.name) {
  
  png(paste0("Nynrin_sgRNA.png"), 1200, 500)
  p<-ggplot(data=df.input, aes(x=mid.point, y=logFC)) +
    geom_bar(stat="identity", fill="darkred", width=50) +
    geom_segment(x=start(gene.coords), y= 0,
                 xend=end(gene.coords),
                 yend= 0,
                 size=3, color="black") +
    theme_minimal() +
    ggtitle(paste0("Nynrin: sgRNAs\n")) +
    ylab("Log Fold Change\n") +
    xlim( start(gene.coords), end(gene.coords) ) +
    ylim( (max(abs(df.final$logFC))+0.25)*-1, max(abs(df.final$logFC))+0.25) +
    theme(plot.title = element_text(size=22),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y   = element_text(size=20),
          #axis.text.x   = element_text(size=20),
          axis.title.y  = element_text(size=20),
          #axis.title.x  = element_text(size=20),
          panel.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.border = element_rect(colour = "black", fill=NA, size=2)) +
    coord_cartesian(clip="off") +
    theme(plot.margin=unit(c(2,2,2,2),"cm"))

  # Add geom segments with for loop
  for (i in 1:nrow(isoform.input)) {
     p <- p + geom_segment(x=isoform.input$start[i], y= 0, 
                           xend=isoform.input$end[i], yend= 0,
                           size=5,
                           color="gray")
  }
  
  print(p)
  
  dev.off()
  
}

plot.gene(df.final, isoform.one)

```






